Void Correction Notes

### 2016-05-10 doc/void_correction.tex
Removed extraneous infinitesimals from the last three integrals.

Current \texttt{Pangloss} code assumes a cosmology of $\Omega_{m,0}=0.25$, $\Omega_{\Lambda,0}=0.75$ (to ensure a flat universe), and $H_0=73$ km$\cdot$s$^{-1}\cdot$Mpc$^{-1}$ (\textcolor{red}{Were these values chosen to match the original Millennium Simulation parameters? A later run that Stefan used for his ray tracing? Known values when Collett et al. was published?}) which leads to an evolving mass density of

@@

Now we must calculate $\rho_h(z)$ given the foreground catalogs of the Millennium Simulation. While each galaxy/halo are on a continuum of redshifts, we snap each halo to one of $N$ redshift planes from $z=0$ to the source redshift of $z_s=1.3857$, each separated by a constant $\Delta z=1.3857/N$ for faster computation (typically, $N=100$ and $\Delta z\approx0.014$). We can approximate $\rho_h$ at each of the $N$ redshift values by summing the halo masses and dividing by the integral the proper volume element $dV_p$ over the redshift range $z\pm \Delta z/2$ and solid angle $d\Omega$ covered by the foreground catalog. Figure \ref{rho_slice} visualizes this process.\\

@@

From Figure \ref{rho_slice}, it is clear that we want a proper volume element that is expressed in terms of the solid angle $d\Omega$ (or really, $d\theta d\phi$) and the redshift $dz$. To do this, we can start with the standard spherical volume element $dV=r^2\sin(\theta)drd\theta d\phi$. However, we must be careful that (1) we are using proper distances, (2) account for the possibility of a curved universe, and (3) use the correct distance measure for $r$ (in this case, the angular diameter distance). From this we can gather that the proper volume element must have the form of

\begin{equation}\label{proper_element}
dV_p=dr_pdA=d_A(z)^2dr_pd\omega=d_A(z)^2\sin(\theta)dr_pd\theta d\phi
\end{equation}
% and $r_a$ is the angular diameter distance 
where $d_A(z)$ is the angular diameter distance and the subscript $p$ denotes proper distance. When the angular diameter distance $d_A(z)$ is multiplied by an angle (i.e. $d_A(z)d\theta$ and $d_A(z)d\phi$), it gives an approximation of the proper distance between two objects at the same redshift. $d_A(z)$ can be expressed in terms of the transverse comoving distance $d_M(z)$,

@@

\begin{equation}\label{temp_solution}
dV_p=d_a(z)^2\sin(\theta)dr_pd\theta d\phi=\left(\frac{d_m(z)}{(1+z)}\right)^2\sin(\theta)dr_pd\theta d\phi=\frac{d_m(z)^2\sin(\theta)}{(1+z)^2}dr_pd\theta d\phi
\end{equation}

@@

\begin{align}
dV_p&=\frac{d_m(z)^2\sin(\theta)}{(1+z)^2}dr_pd\theta d\phi\nonumber\\
&=\frac{d_m(z)^2\sin(\theta)}{(1+z)^2}\left(\frac{d_H}{E(z)(1+z)}dz\right)d\theta d\phi\nonumber\\
&=\frac{d_Hd_M(z)^2\sin(\theta)}{E(z)(1+z)^3}dzd\theta d\phi.
\end{align}

We can calculate the proper volume of a given grid slice at redshift $z$ in a foreground catalog centered at ($\theta,\phi$) by integrating the above volume element over the $\Delta z$, $\Delta\theta$, and $\Delta\phi$ range of the redshift slice:

\begin{equation}
V(z)=\int_{V_z}dV_p=d_H\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}}\int_{\theta-\frac{\Delta\theta}{2}}^{\theta+\frac{\Delta\theta}{2}}\int_{\phi-\frac{\Delta\phi}{2}}^{\phi+\frac{\Delta\phi}{2}} \frac{d_M(z^\prime)^2\sin(\theta^\prime)}{E(z^\prime)(1+z^\prime)^3}dz^\prime d\theta^\prime d\phi^\prime.
\end{equation}

Integrating over $d\theta^\prime$ and $d\phi^\prime$ and using the trig identity ${\cos(u)-\cos(v)=2\sin\left(\frac{u+v}{2}\right)\sin\left(\frac{u-v}{2}\right)}$ gives

\begin{equation}\label{volume_final}
V(z)=2d_H\Delta\phi\sin(\theta)\sin(\Delta\theta/2)\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2\sin(\theta^\prime)}{E(z^\prime)(1+z^\prime)^3}dz^\prime d\theta^\prime d\phi^\prime
\end{equation}

which is our final expression for the proper volume.\\

@@

\begin{equation}\label{rho_h}
\rho_h(z)=\frac{\sum_i M_{i,z}}{V(z)}=\frac{\sum_i M_{i,z}}{2d_H\Delta\phi\sin(\theta)\sin(\Delta\theta/2)\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2\sin(\theta^\prime)}{E(z^\prime)(1+z^\prime)^3}dz^\prime d\theta^\prime d\phi^\prime}.
\end{equation}

@@

\begin{align}\label{smooth_component} 
\rho_s(z)&=\rho_m(z)-\rho_h(z)\nonumber\\
&=\left(\frac{3H_0^2}{8\pi G}\,\Omega_{m,0}\right)(1+z)^3 - \frac{\sum_i M_{i,z}}{2d_H\Delta\phi\sin(\theta)\sin(\Delta\theta/2)\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2\sin(\theta^\prime)}{E(z^\prime)(1+z^\prime)^3}dz^\prime d\theta^\prime d\phi^\prime}
\end{align}

>>>>

Current \texttt{Pangloss} code assumes a cosmology of $\Omega_{m,0}=0.25$, $\Omega_{\Lambda,0}=0.75$ (to ensure a flat universe), and $H_0=73$ km$\cdot$s$^{-1}\cdot$Mpc$^{-1}$ which leads to an evolving mass density of

@@

Now we must calculate $\rho_h(z)$ given the foreground catalogs of the Millennium Simulation. While each galaxy/halo are on a continuum of redshifts, we snap each halo to one of $N$ redshift planes from $z=0$ to the source redshift of $z_s=1.3857$, each separated by a constant $\Delta z=1.3857/N$ for faster computation (typically, $N=100$ and $\Delta z\approx0.014$). We can approximate $\rho_h$ at each of the $N$ redshift values by summing the halo masses and dividing by the integral the proper volume element $dV_p$ over the redshift range $z\pm \Delta z/2$ and solid angle $d\Omega$ covered by the lightcone. Figure \ref{rho_slice} visualizes this process.\\

@@

From Figure \ref{rho_slice}, it is clear that we want a proper volume element that is expressed in terms of the solid angle $d\Omega$ and the redshift $dz$. To do this, we can start with the standard spherical volume element $dV=r^2drd\Omega$. However, we must be careful that (1) we are using proper distances, (2) account for the possibility of a curved universe, and (3) use the correct distance measure for $r$ (in this case, the angular diameter distance). From this we can gather that the proper volume element must have the form of

\begin{equation}\label{proper_element}
dV_p=d_A(z)^2dr_pd\Omega
\end{equation}
% and $r_a$ is the angular diameter distance 
where $d_A(z)$ is the angular diameter distance and the subscript $p$ denotes proper distance. When the angular diameter distance $d_A(z)$ is multiplied by an angle (i.e. $d_A(z)d\theta$ and $d_A(z)d\phi$), it gives an approximation of the proper distance between two objects at the same redshift. $d_A(z)$ can be expressed in terms of the transverse comoving distance $d_M(z)$,

@@

\begin{equation}\label{temp_solution}
dV_p=d_a(z)^2dr_pd\Omega=\left(\frac{d_m(z)}{(1+z)}\right)^2dr_pd\Omega=\frac{d_m(z)^2}{(1+z)^2}dr_pd\Omega
\end{equation}

@@

\begin{align}
dV_p&=\frac{d_m(z)^2}{(1+z)^2}dr_pd\Omega\nonumber\\
&=\frac{d_m(z)^2}{(1+z)^2}\left(\frac{d_H}{E(z)(1+z)}dz\right)d\Omega\nonumber\\
&=\frac{d_Hd_M(z)^2}{E(z)(1+z)^3}dzd\Omega.
\end{align}

We can calculate the proper volume of the lightcone a given grid slice at redshift $z$ in a foreground catalog centered at ($\theta,\phi$) by integrating the above volume element over the solid angle $d\Omega$ and $\Delta z$ range of the redshift slice:

\begin{equation}
V(z)=\int_{V_z}dV_p=d_H\int_\Omega\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2}{E(z^\prime)(1+z^\prime)^3}dz^\prime d\Omega.
\end{equation}

Integrating over $d\Omega$ results in the solid angle of a cone given by $\Omega=2\pi(1-\cos\theta)$, so the previous equation becomes

\begin{equation}\label{volume_final}
V(z)=2\pi(1-\cos\theta)d_H\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2}{E(z^\prime)(1+z^\prime)^3}dz^\prime
\end{equation}

which is our final equation for the proper volume.\\

@@

\begin{equation}\label{rho_h}
\rho_h(z)=\frac{\sum_i M_{i,z}}{V(z)}=\frac{\sum_i M_{i,z}}{2\pi(1-\cos\theta)d_H\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2}{E(z^\prime)(1+z^\prime)^3}dz^\prime}.
\end{equation}

@@

\begin{align}\label{smooth_component} 
\rho_s(z)&=\rho_m(z)-\rho_h(z)\nonumber\\
&=\left(\frac{3H_0^2}{8\pi G}\,\Omega_{m,0}\right)(1+z)^3 - \frac{\sum_i M_{i,z}}{2\pi(1-\cos\theta)d_H\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \frac{d_M(z^\prime)^2}{E(z^\prime)(1+z^\prime)^3}dz^\prime}
\end{align}
